---
title: "Analysis of large BE data set - without using iBARs"
author: "Magdalena Strauss"
output:
  html_document:
    number_sections: yes
    keep_md: yes
---

This script processing the large single-cell base editor screen with DNA-sequencing counterpart. 

# Setup and loading SingleCellExperiment with gRNAs assigned

```{r,setup}
library(bluster)
source("../core_functions.R")
library("batchelor")
library("scater")
library(scran)
library(dplyr)
library(igraph)
library(RColorBrewer)
library(pheatmap)
library(Seurat)
library(MASS)
library(clusterProfiler)
library(DOSE)
library(ReactomePA)
library(org.Hs.eg.db)
library(energy)
library(ggrepel)
library(uwot)
set.seed(4444)

knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE,
    message = FALSE,
    cache=FALSE,
    dev = c("pdf"),
    dpi=300
)

sce_file <- "sce_bigBE_gRNA_iBAR.rds"
sce_save_file <- "sce_bigBE_gRNA_analysed.rds"
folder_cellranger <-"/nfs/research/marioni/magda/scJAK1_August_2022_cellranger/bigBE_1/outs"
markers_file <- paste0("markers_bigBE.csv")
markers_file_full <- paste0("markers_bigBE_full.csv")

colours_consequence <-  c("stop codon"="darkred","missense"="red","splice variant"="orange","promoter"="yellow","UTR"="purple","het"="darkblue","WT"="#000000","intron"="green","synonymous"="#ADD8E6")
```

Reading in SingleCellExperiment.

```{r}
sce <- readRDS(sce_file)
features <- read.delim(paste0(folder_cellranger,"/filtered_feature_bc_matrix/features.tsv.gz"),header=FALSE,stringsAsFactors = FALSE)
rowData(sce) <- features[match(rownames(sce),features$V1),]
```

Subsetting to cells with a unique gRNA and computing UMAP.

```{r}
sce <- sce[,!(is.na(sce$gRNA))]
sce <- sce[,!(grepl("-",sce$gRNA))]
```

Adding meta data

```{r}
JAK1_mutagenesis <- read.table("../MissionBio/mutation_file_Coelho_2023.csv",header=TRUE,sep=",")
sce$zscore_FACS <- JAK1_mutagenesis$zscore_FACS[match(sce$gRNA,JAK1_mutagenesis$sgRNA_ID)]
sce$predicted_consequence <- JAK1_mutagenesis$Consequence[match(sce$gRNA,JAK1_mutagenesis$sgRNA_ID)]
sce$zscore_proliferation <- JAK1_mutagenesis$zscore_proliferation[match(sce$gRNA,JAK1_mutagenesis$sgRNA_ID)]
feature_ref <- read.table("feature_ref.csv",header=TRUE,sep=",")
sce$target_gene <- feature_ref$target_gene_name[match(sce$gRNA,feature_ref$id)]
```

# Find gRNAs with similar transcriptional impact using lower dimensional representation

## Identifying a group of impacted genes 

Differential expression analysis for each gRNA compared to the nontargeting control identifies impacted genes. 

```{r}
sce$gRNA[sce$target_gene=="Non-Targeting"] <- "NT"
gRNAs <- names(table(sce$gRNA[table(sce$gRNA) >= 10]))
gRNAs <- setdiff(gRNAs,"NT")
```

Perform a Wilcoxon test for each gRNA with at least 10 cells identify differential expressino

```{r, eval=FALSE}
markers <- list()
for (j in 1:length(gRNAs)){
  sce_temp <- sce[,sce$gRNA%in%c("NT",gRNAs[j])]
  temp <- findMarkers(sce_temp,groups=sce_temp$gRNA,test.type="wilcox")
  temp <- temp[[gRNAs[j]]]
  temp$gRNA <- gRNAs[j]
  temp$target_gene <- sce_temp$gene[sce_temp$gRNA==gRNAs[j]][1]
  temp$DE_gene <- rownames(temp)
  temp$DE_gene_name <- rowData(sce_temp)$V2[match(rownames(temp),rownames(sce))]
  temp$target_gene <- sce_temp$target_gene[which(sce_temp$gRNA==gRNAs[j])[1]]
  markers[[j]] <- temp[order(rownames(temp)),]
}
FDRs <- lapply(markers,function(x) x$FDR)
FDRs <- do.call(cbind,FDRs)
FDRs <- apply(FDRs,1,p.adjust)
for (j in 1:length(markers)){
  markers[[j]]$FDR <- FDRs[j,]
}
markers <- do.call(rbind,markers)
markers$rho <- abs(markers$summary.AUC-0.5)
write.table(markers[markers$FDR<0.1,],file=markers_file,sep=",",col.names = TRUE,row.names=FALSE)
write.table(markers,file=markers_file_full,sep=",",col.names = TRUE,row.names=FALSE)
```

Finding a basis of genes: we combine genes that are differentially expressed for at least of the gRNAs and the JAK-STAT pathway genes. 

```{r }
JAK_STAT_pathway_genes <- read.table("../JAK_STAT_pathway_genes.txt")$V1
JAK_STAT_pathway_genes <- features$V1[match(JAK_STAT_pathway_genes,features$V2)]
markers <-  read.table(markers_file,sep=",",header=TRUE)
basis_of_genes <- unique(c(JAK_STAT_pathway_genes,markers$DE_gene))
```

The basis of genes has `r length(basis_of_genes)` genes. 

PCA is now performed on this basis of genes. 

```{r }
reducedDims(sce)$PCA_basis <- calculatePCA(sce[rowData(sce)$V1%in%basis_of_genes,],ncomponents=50,scale=TRUE)
```

# UMAPs illustrating predicted consequences and previous bulk-based analysis

First we compute UMAP coordinates. 

```{r}
sce_UMAP <- uwot::umap(reducedDims(sce)$PCA_basis,min_dist=0.5)
reducedDims(sce)$UMAP <- sce_UMAP
```

Coloured by the FACS-based z-score obtained from Coelho, M. A. et al. Base editing screens map mutations affecting interferon-γ signaling in cancer. Cancer Cell 41, 288–303.e6 (2023).


```{r umap_zscore_FACS}
plotReducedDim(sce,colour_by = "zscore_FACS",dimred="UMAP")+ theme(legend.position = "bottom")
```

Plot of UMAP coloured by proliferation z-score (Coelho et al. 2023). 

```{r umap_zscore_proliferation}
plotReducedDim(sce,colour_by = "zscore_proliferation",dimred="UMAP")+ theme(legend.position = "bottom")
```

UMAP coloured by the consequence of the mutation predicted by Coelho et al. (2023). 

```{r umap_predicted_consequence}
plotReducedDim(sce,colour_by = "predicted_consequence",dimred="UMAP") +
  scale_color_manual(values=colours_consequence)+
  labs(color="predicted consequence")
```
Plotting the predicted consequences individually

```{r umap_predicted_consequence_ind}
sce$predicted_consequence[is.na(sce$predicted_consequence)] <- "WT"
for (j in 1:length(unique(sce$predicted_consequence))){
  sce_temp <- cbind(sce[,sce$predicted_consequence!=unique(sce$predicted_consequence)[j]],
                   sce[,sce$predicted_consequence==unique(sce$predicted_consequence)[j]] )
sce_temp$predicted_consequence[sce_temp$predicted_consequence!=unique(sce$predicted_consequence)[j]] <- NA
print(plotReducedDim(sce_temp,colour_by = "predicted_consequence",dimred="UMAP") +
  scale_color_manual(values=colours_consequence,na.value="lightgrey")+
  labs(color="predicted consequence")+ theme_classic(base_size=12)+ theme(legend.position = "bottom")+guides(color=guide_legend(nrow=2,override.aes = list(size=3))))}
```


# Clustering 

Louvain clustering is performed in the PC space. 

```{r}
g <- clusterCells(sce, use.dimred="PCA_basis", 
    BLUSPARAM=NNGraphParam(cluster.fun="louvain"))
sce$cluster <- g
sce$cluster <- as.factor(sce$cluster)
```

Plotting the clusters on the UMAP

```{r umap_clusters}
plotReducedDim(sce,colour_by = "cluster",dimred="UMAP",point_size=0.3)+ theme_classic(base_size=12)+ theme(legend.position = "bottom")+guides(color=guide_legend(nrow=2,override.aes = list(size=3)))
```

Distribution of predicted mutational consequence across clusters

```{r cluster_consequence}
sce$predicted_consequence[is.na(sce$predicted_consequence)] <- "WT"
print(ggplot(mapping=aes(x=sce$cluster,fill=sce$predicted_consequence)) +geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..])) +
     ylab("") +labs(fill="") + xlab("cluster") + theme_bw(base_size=15)+ggtitle("Predicted consequence")+
  scale_fill_manual(values=colours_consequence) + theme(legend.position = "bottom") + guides(fill = guide_legend(nrow = 4))+ylab("number of cells"))

```

Proportion of control gRNAs in each cluster

```{r proportion_of_NT_gRNA}
cluster_gRNA_count_matrix <- reshape2::acast(data.frame(gRNA=sce$gRNA,cluster=sce$cluster),gRNA~cluster)
prop_NT_gRNA <- cluster_gRNA_count_matrix["NT",]/colSums(cluster_gRNA_count_matrix)
ggplot(mapping=aes(x=names(prop_NT_gRNA),y=prop_NT_gRNA)) + geom_bar(stat="identity")+theme_classic(base_size=22)+xlab("cluster")+
  ylab("proportion NT-gRNA")
```

Combining clusters with similar proportion of non-targeting gRNAs

```{r meta_clusters_2}
meta_clusters_2 <- kmeans(prop_NT_gRNA,centers=2)$cluster
meta_clusters_2[meta_clusters_2==1] <- "LoF"
meta_clusters_2[meta_clusters_2==2] <- "not_LoF"
sce$meta_cluster_2 <- meta_clusters_2[sce$cluster]
xx <- table(sce$meta_cluster_2[sce$gRNA=="NT"])
if (xx["LoF"] > xx["not_LoF"]){
  temp <- sce$meta_cluster_2
  sce$meta_cluster_2[temp=="LoF"] <- "not_LoF"
  sce$meta_cluster_2[temp=="not_LoF"] <- "LoF"
}

plotReducedDim(sce,colour_by = "meta_cluster_2",dimred="UMAP",point_size=0.3)+guides(color=guide_legend(nrow=2,override.aes = list(size=3))) + scale_color_manual(values=c("LoF"="purple","not_LoF"="#5C4033"))+ theme(legend.position = "bottom")+guides(color=guide_legend(nrow=2,override.aes = list(size=3)))
```

Distribution of predicted consequence over the two meta-clusters

```{r meta_cluster_consequence}
print(ggplot(mapping=aes(x=sce$meta_cluster_2,fill=sce$predicted_consequence)) +geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..])) +
     ylab("") +labs(fill="") + xlab("meta cluster") + theme_bw(base_size=15)+ggtitle("Predicted consequence")+
  scale_fill_manual(values=colours_consequence) + theme(legend.position = "bottom") + guides(fill = guide_legend(nrow = 4))+ylab("number of cells"))

```

# Differential expression analysis 

Comparing cells with gRNAs with more than 70% of cells, and at least 3 cells, in meta-cluster LoF to non-targeting gRNAs.

```{r DE_LoF}
df_gRNA_meta_cluster <- data.frame(gRNA=sce$gRNA,meta_cluster=sce$meta_cluster_2)
count_gRNA_meta_cluster <- df_gRNA_meta_cluster %>% dplyr::group_by_all() %>% dplyr::count()
count_gRNA_meta_cluster <- count_gRNA_meta_cluster[order(count_gRNA_meta_cluster$gRNA),]
count_gRNA_meta_cluster$freq <- NA
for (j in 1:nrow(count_gRNA_meta_cluster)){
  count_gRNA_meta_cluster$freq[j] <- count_gRNA_meta_cluster$n[j]/sum(count_gRNA_meta_cluster$n[count_gRNA_meta_cluster$gRNA==count_gRNA_meta_cluster$gRNA[j]])
}
xx <- count_gRNA_meta_cluster$n>=3
xx[count_gRNA_meta_cluster$freq<0.7] <- F
xx[count_gRNA_meta_cluster$meta_cluster!="LoF"] <- F
count_gRNA_meta_cluster_LoF <- count_gRNA_meta_cluster[xx,]
gRNAs_LoF <- count_gRNA_meta_cluster_LoF$gRNA
sce_temp <- sce[,sce$gRNA%in%c(gRNAs_LoF,"NT")]
sce_temp$LoF <- sce_temp$gRNA != "NT"
markers_JAK_STAT_LoF <- findMarkers(sce_temp[JAK_STAT_pathway_genes,],group=sce_temp$LoF,test.type="wilcox")[["TRUE"]]
markers_JAK_STAT_LoF$gene_name <- features$V2[match(rownames(markers_JAK_STAT_LoF),features$V1)]
df <- markers_JAK_STAT_LoF[,c(3,4,6)]
df <- as.data.frame(df)
ggplot(df, aes(x=summary.AUC, y=-log10(FDR))) +
    geom_point(aes(color="grey", size=1, alpha=0.4)) +
    ggtitle('Volcano Plot - JAK-STAT pathway') +
    labs(y=expression('-Log'[10]*' FDR'), x=expression('AUC')) +
    theme_classic(base_size=20) +
    theme(legend.position="none", plot.title = element_text(size = rel(1), hjust = 0.5))+
    geom_text_repel(data=df[df$FDR<0.1,],aes(x = summary.AUC, y = -log10(FDR),label=gene_name),max.overlaps=100)+
    geom_point(data=df[df$FDR<0.1,],aes(x = summary.AUC, y = -log10(FDR),color="red", size=1,alpha=0.4))+
    scale_color_manual(values=c('gray','red'))+ scale_y_continuous(trans=scales::pseudo_log_trans(sigma=1,base = 1.05))+
  geom_vline(xintercept = 0.5)+geom_hline(yintercept = 1)
```

Plotting the proportions of gRNAs in the LoF meta-cluster for gRNAs with at least 3 cells assigned. 

```{r proportion_meta_cluster}
gRNA_meta_cluster_count <- reshape2::acast(df_gRNA_meta_cluster,gRNA~meta_cluster)
gRNA_meta_cluster_count <- gRNA_meta_cluster_count[rowSums(gRNA_meta_cluster_count)>=3,]
gRNA_meta_cluster_freq <- t(apply(gRNA_meta_cluster_count,1,function(x) return(x/sum(x))))
ggplot(mapping=aes(x=gRNA_meta_cluster_freq)) + geom_histogram()+theme_classic(base_size=20)+
  xlab("proportion in LoF-cluster") + ylab("number of gRNAs")
```


```{r}
saveRDS(sce,file=sce_save_file)
saveRDS(JAK1_mutagenesis,file="JAK1_mutatgenesis_extended.rds")
```
