---
title: "QC for mRNA data - large BE data set"
author: "Magdalena Strauss"
output:
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
---

This notebook performs outlier-based QC for the non-genotyped large base editor screen, removing low outliers for the total count, low outliers for the number of detected features and high outliers for the percentage of counts from mitochondrial genes.

```{r}
# knitr::opts_chunk$set(
#     echo = FALSE,
#     message = FALSE,
#     warning = FALSE,
#     message = FALSE,
#     dev = "pdf",
#     dpi=300
# )

source("../core_functions.R")

library(Matrix)
library(ggplot2)
library(AnnotationHub)
library(AnnotationDbi)
library(scran)
library(scater)
library(igraph)
library(gridExtra)

folders_cellranger <- paste0("/nfs/research/marioni/magda/scJAK1_August_2022_cellranger/bigBE_",sapply(1:4,toString),"/outs")
file_name_mRNA_matrices <- "mRNA_matrices_bigBE.rds"
file_name_mRNA_matrices_QC <- "mRNA_matrices_QC_bigBE.rds"
file_name_sces <- "sce_list_bigBE.rds"
ids <- c("sample1","sample2","sample3","sample4")
```

```{r}
mRNA_matrices <- list()
total_cell_number <- 0
for (k in 1:length(folders_cellranger)){
 matrix_temp <- read_in_filtered_matrix(folders_cellranger[k])
  mRNA_matrices[[k]] <- matrix_temp[grepl("ENS",rownames(matrix_temp)),]
  total_cell_number <- total_cell_number + ncol(mRNA_matrices[[k]])
}
names(mRNA_matrices) <- ids
saveRDS(mRNA_matrices,file_name_mRNA_matrices)
```

The total number of cells before QC is `r total_cell_number`.

```{r}
discard <- list()
mRNA_matrices_QC <- list()
total_cell_number_after_QC <- 0
for (j in 1:length(mRNA_matrices)){
  discard[[j]] <- QC_mRNA_outlier(mRNA_matrices[[j]],file_name = paste0("Sample_",ids[j]))
  mRNA_matrices_QC[[j]] <- mRNA_matrices[[j]][,!(discard[[j]])]
  total_cell_number_after_QC <- total_cell_number_after_QC + ncol(mRNA_matrices_QC[[j]])
}
names(mRNA_matrices_QC) <- ids
saveRDS(mRNA_matrices_QC,file=file_name_mRNA_matrices_QC)
```

The total number of cells after QC is `r  total_cell_number_after_QC`.

```{r}
set.seed(42)

sce_list <- list()
for (j in 1:4){
  sce_list[[j]] <- SingleCellExperiment(
    assays = list(counts = mRNA_matrices_QC[[j]]), colData = colnames(mRNA_matrices_QC[[j]]))
  clusts = as.numeric(quickCluster(sce_list[[j]], method = "igraph", min.size = 100))
  sce_list[[j]] = computeSumFactors(sce_list[[j]], clusters = clusts)
  sce_list[[j]] <- logNormCounts(sce_list[[j]])
}
names(sce_list) <- ids
saveRDS(sce_list,file=file_name_sces)
```


