---
title: "Genotype assignment per barcode"
output:
  html_document:
    number_sections: yes
    keep_md: yes
---

This script processes per-cell genotype calls using iBAR and puroR-barcode information. The output is summarised genotype information per barcode. Further, consequence and impact of genotypes are predicted using VEP (W. McLaren et al. The ensembl variant effect predictor. Genome Biol., 17 (2016)) and actual genotype information. 

```{r,setup}
knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE,
    message = FALSE,
    dev = c("pdf"),
    dpi=300
)
source("core_functions_MissionBio.R")
library(dplyr)
```

Reading in the per-cell genotype information. 

```{r}
df_GT <- readRDS("genotypes_MB_per_cell.rds")
puroBC.keep <- names(table(df_GT$puroBC_gRNA_iBAR))[table(df_GT$puroBC_gRNA_iBAR)>= 3]
df_GT_min_cells <- df_GT[df_GT$puroBC_gRNA_iBAR%in%puroBC.keep,]
df_GT_min_cells <- df_GT_min_cells[!(grepl("NA",df_GT_min_cells$puroBC_gRNA_iBAR)),]
dim(df_GT_min_cells)
```

Checking if any puro-barcodes are associated with more than one gRNA-iBAR combination. 

```{r}
count_puroBC_iBAR <- df_GT_min_cells[,3:4] %>% dplyr::group_by_all() %>% dplyr::count()
puroBC_dup <- names(table(count_puroBC_iBAR$puroBC))[table(count_puroBC_iBAR$puroBC) > 1]
df_GT_min_cells <- df_GT_min_cells[!(df_GT_min_cells$puroBC%in%puroBC_dup),]
```

Now we assign mutations to puro-barcodes.

```{r}
GT_puro_BC <-  assign_mutations_to_puro_barcodes(df_GT_min_cell_number=df_GT_min_cells, column_name_barcode="puroBC")
GT_puro_BC <- GT_puro_BC[unlist(lapply(GT_puro_BC,function(x) x!=""))]
df_GT_sub <- df_GT[df_GT$puroBC%in%names(GT_puro_BC),]
barcode_gRNA <- data.frame(cell=df_GT_sub$cell,puroBC=df_GT_sub$puroBC,gRNA_iBAR=paste0(df_GT_sub$gRNA,
                                                                                          "-",df_GT_sub$iBAR))
barcode_gRNA_count <- barcode_gRNA[,2:3] %>% dplyr::group_by_all() %>% dplyr::count()
barcode_gRNA_count <- barcode_gRNA_count[!(grepl("NA",barcode_gRNA_count$gRNA_iBAR)),]
barcode_gRNA_count <- barcode_gRNA_count[barcode_gRNA_count$n>1,]
barcode_gRNA_count$GT <- GT_puro_BC[barcode_gRNA_count$puroBC]
df_GT_puro_iBAR <- barcode_gRNA_count[,c(1,2,4)] 
df_GT_puro_iBAR$any_homozygous <- grepl("1/1/1",df_GT_puro_iBAR$GT)


df_GT_puro_iBAR$gRNA <- sapply(df_GT_puro_iBAR$gRNA_iBAR,function(x) strsplit(x,"-")[[1]][1])
gRNA_info <- data.table::fread("gRNA_info_coordinates.csv")
u_gRNA <- unique(df_GT_puro_iBAR$gRNA)
u_gRNA <- u_gRNA[!(is.na(u_gRNA))]
df_GT_puro_iBAR$gene <- rep("",nrow(df_GT_puro_iBAR))
for (j in 1:length(u_gRNA)){
  df_GT_puro_iBAR$gene[which(df_GT_puro_iBAR$gRNA == u_gRNA[j])] <- gRNA_info$Gene[which(gRNA_info$sgRNA_ID == u_gRNA[j])]
}

df_GT_puro_iBAR$start <- NA 
df_GT_puro_iBAR$strand <- NA 
df_GT_puro_iBAR$end <- NA 
df_GT_puro_iBAR$exonic <- NA
for (j in 1:length(u_gRNA)){
  df_GT_puro_iBAR$strand[df_GT_puro_iBAR$gRNA %in% u_gRNA[j]] <- unique(gRNA_info$strand[gRNA_info$sgRNA_ID == u_gRNA[j]]) 
} 
 for (j in 1:length(u_gRNA)){
  df_GT_puro_iBAR$start[df_GT_puro_iBAR$gRNA %in% u_gRNA[j]] <- unique(gRNA_info$start[gRNA_info$sgRNA_ID == u_gRNA[j]]) }
for (j in 1:length(u_gRNA)){
  df_GT_puro_iBAR$end[df_GT_puro_iBAR$gRNA %in% u_gRNA[j]] <- unique(gRNA_info$end[gRNA_info$sgRNA_ID == u_gRNA[j]]) }
for (j in 1:length(u_gRNA)){
  df_GT_puro_iBAR$exonic[df_GT_puro_iBAR$gRNA %in% u_gRNA[j]] <- unique(gRNA_info$exonic[gRNA_info$sgRNA_ID == u_gRNA[j]]) }
```

Removing barcodes with additional non C->T edits at the same location as a C -> T edit.

```{r}
df_GT_puro_iBAR <- df_GT_puro_iBAR[!(grepl("/2",df_GT_puro_iBAR$GT)),]
```

The number of puro-barcodes after this filtering is the following:

```{r}
length(unique(df_GT_puro_iBAR$puroBC))
```

The mutations are converted to an input format for VEP (W. McLaren et al. The ensembl variant effect predictor. Genome Biol., 17 (2016)). 

```{r}
mutations_all <- unlist(sapply(df_GT_puro_iBAR$GT,function(x) strsplit(x,";")))
names(mutations_all) <- NULL
mutations_all <- unique(mutations_all)
positions_all <- sapply(mutations_all,function(x) strsplit(x,"-")[[1]][c(1,3,4)])
dim(positions_all) <- c(3,length(positions_all)/3)
ref <- positions_all[2,]
alt <- ref
alt[ref=="C"] <- "T"
alt[ref=="G"] <- "A"
df_VEP <- data.frame(chrom=rep(1,ncol(positions_all)),start=positions_all[1,],end=positions_all[1,],allele=paste0(ref,"/",alt),
                     strand="+")
df_VEP <- distinct(df_VEP)
write.table(df_VEP,"VEP_input.txt",quote=FALSE,col.names=FALSE,row.names=FALSE)
```

VEP output is loaded and added to the data frame for the homozygous mutations. 

```{r}
VEP<- data.table::fread("VEP_MB2.txt" ,header=TRUE)
VEP$mutation <- sapply(VEP$`#Uploaded_variation`,function(x) paste0(strsplit(x,"_")[[1]][2],"-",strsplit(x,"_")[[1]][3]))
VEP <- VEP[VEP$MANE_SELECT!="-",]
VEP$gene <- VEP$SYMBOL
VEP <- VEP[VEP$gene=="JAK1",]
VEP <- VEP[,c("mutation","Consequence","IMPACT","Existing_variation","gene","SIFT","PolyPhen","MaxEntScan_diff")]
VEP$Existing_variation <- sapply(VEP$Existing_variation,function(x) gsub(",","-",x))
rownames(VEP) <- NULL
VEP <- dplyr::distinct(VEP)
```

We classify mutations as in Coelho et al. Base editing screens map mutations affecting IFNg signalling in cancer. Cancer Cell (2023), see https://github.com/MatthewACoelho/Base_Editing_Screens/blob/main/ABE8e_080522.Rmd

```{r}
# https://github.com/MatthewACoelho/Base_Editing_Screens/blob/main/ABE8e_080522.Rmd
Consequence <- VEP$Consequence
Consequence <- sapply(Consequence,function(x) gsub(",","_",x))
Consequence = ifelse(Consequence == "missense_variant", "missense", ifelse(Consequence == "missense_variant_splice_region_variant", "splice variant", 
                    ifelse(Consequence == "splice_region_variant", "splice variant", 
                    ifelse(Consequence == "stop_retained_variant", "synonymous",
                    ifelse(Consequence == "synonymous_variant", "synonymous", 
                    ifelse(Consequence == "splice_region_variant_synonymous_variant", "splice variant",
                     ifelse(Consequence == "splice_region_variant,synonymous_variant" ,"splice variant",
                    ifelse(Consequence == "splice_acceptor_variant", "splice variant",
                    ifelse(Consequence == "splice_donor_variant", "splice variant",
                    ifelse(Consequence == "splice_region_variant_intron_variant", "splice variant",
                    ifelse(Consequence == "splice_polypyrimidine_tract_variant", "splice variant",
                    ifelse(Consequence == "splice_polypyrimidine_tract_variant,intron_variant", "splice variant",
                    ifelse(Consequence == "splice_donor_region_variant_intron_variant", "splice variant",
                    ifelse(Consequence == "splice_polypyrimidine_tract_variant_intron_variant", "splice variant",
                    ifelse(Consequence == "splice_polypyrimidine_tract_variant_splice_region_variant_intron_variant", "splice variant",
                    ifelse(Consequence == "splice_donor_5th_base_variant_intron_variant", "splice variant",
                    ifelse(Consequence == "downstream_gene_variant", "UTR",
                    ifelse(Consequence == "stop_gained_splice_region_variant", "stop codon",
                    ifelse(Consequence == "start_lost", "start lost",
                    ifelse(Consequence == "stop_gained_start_lost", "stop codon",
                    ifelse(Consequence == "upstream_gene_variant", "promoter",
                    ifelse(Consequence == "intron_variant", "intron",
                    ifelse(Consequence == "5_prime_UTR_variant", "5'UTR",
                    ifelse(Consequence == "3_prime_UTR_variant", "3'UTR",
                    ifelse(Consequence == "stop_gained", "stop codon", Consequence)))))))))))))))))))))))))
VEP$Consequence <- Consequence
#Label a mutation 'promoter' if it is in the promotor region of JAK1 as identified in Coelho et. al (2023)
VEP_positions <- as.double(sapply(VEP$mutation,function(x) strsplit(x,"-")[[1]][1]))
VEP$Consequence[VEP_positions >= 64964978 &  VEP_positions <= 64967543] <- "promoter"
```

```{r}
df_GT_puro_iBAR_edited <- df_GT_puro_iBAR[!(df_GT_puro_iBAR$GT%in%c("WT","")),]
mutation_list <- sapply(df_GT_puro_iBAR_edited$GT,function(x) strsplit(x,";"))
df_GT_puro_iBAR$impact <- NA
df_GT_puro_iBAR$consequence <- NA
df_GT_puro_iBAR$existing_variation <- NA
df_GT_puro_iBAR$homozygous_edits <- NA
df_GT_puro_iBAR$SIFT <- NA
df_GT_puro_iBAR$PolyPhen <- NA
df_GT_puro_iBAR$MaxEntScan_diff <- NA
for (j in 1:length(mutation_list)){
  temp <- sapply(mutation_list[[j]],function(x) strsplit(x,"-")[[1]][c(1,3,4)])
  type_temp <- sapply(mutation_list[[j]],function(x) strsplit(x,"-")[[1]][2])
  type <- rep("het",length(type_temp))
  type[type_temp=="1/1/1"] <- "hom"
  ref <- temp[2,]
  alt <- ref
  alt[ref=="C"] <- "T"
  alt[ref=="G"] <- "A"
  df_VEP_temp <- paste0(temp[1,],"-",ref,"/",alt)
  existing_variation_temp <-  VEP$Existing_variation[match(df_VEP_temp,VEP$mutation)]
  if (length(existing_variation_temp) >1){
      existing_variation_temp <- sort(existing_variation_temp)
      v <- existing_variation_temp[1]
      for (k in 2:length(existing_variation_temp)){
        v <- paste0(v,";",existing_variation_temp[k])
      }
       existing_variation_temp <- v
  }
  df_GT_puro_iBAR$existing_variation[df_GT_puro_iBAR$puroBC==names(mutation_list)[j]] <- existing_variation_temp

  xx <- type=="hom"
  if (sum(xx) > 0){
    impact_temp <-VEP$IMPACT[match(df_VEP_temp[xx],VEP$mutation)]
    impact_temp <- unique(impact_temp)
    if (length(impact_temp) >1){
      impact_temp <- sort(impact_temp)
      v <- impact_temp[1]
      for (k in 2:length(impact_temp)){
        v <- paste0(v,"-",impact_temp[k])
      }
      impact_temp <- v
    }
    if (grepl("HIGH",impact_temp)){
      impact_temp <- "HIGH"
    }
   df_GT_puro_iBAR$impact[df_GT_puro_iBAR$puroBC==names(mutation_list)[j]] <- impact_temp
    
    consequence_temp <- VEP$Consequence[match(df_VEP_temp[xx],VEP$mutation)]
    if (length(consequence_temp) >1){
      consequence_temp <- sort(consequence_temp)
      v <- consequence_temp[1]
      for (k in 2:length(consequence_temp)){
        v <- paste0(v,";",consequence_temp[k])
      }
       consequence_temp <- v
    }
    df_GT_puro_iBAR$consequence[df_GT_puro_iBAR$puroBC==names(mutation_list)[j]] <- consequence_temp
    
    SIFT_temp <- VEP$SIFT[match(df_VEP_temp[xx],VEP$mutation)]
    if (length(SIFT_temp) >1){
      SIFT_temp <- sort(SIFT_temp)
      v <- SIFT_temp[1]
      for (k in 2:length(SIFT_temp)){
        v <- paste0(v,";",SIFT_temp[k])
      }
       SIFT_temp <- v
    }
    df_GT_puro_iBAR$SIFT[df_GT_puro_iBAR$puroBC==names(mutation_list)[j]] <- SIFT_temp
    
    PolyPhen_temp <- VEP$PolyPhen[match(df_VEP_temp[xx],VEP$mutation)]
    if (length(PolyPhen_temp) >1){
      PolyPhen_temp <- sort(PolyPhen_temp)
      v <- PolyPhen_temp[1]
      for (k in 2:length(PolyPhen_temp)){
        v <- paste0(v,";",PolyPhen_temp[k])
      }
       PolyPhen_temp <- v
    }
    df_GT_puro_iBAR$PolyPhen[df_GT_puro_iBAR$puroBC==names(mutation_list)[j]] <- PolyPhen_temp
    
    MaxEntScan_diff_temp <- as.double(VEP$MaxEntScan_diff[match(df_VEP_temp[xx],VEP$mutation)])
    MaxEntScan_diff_temp <- MaxEntScan_diff_temp[!(is.na(MaxEntScan_diff_temp))]
    if (length(MaxEntScan_diff_temp)>0){
    df_GT_puro_iBAR$MaxEntScan_diff[df_GT_puro_iBAR$puroBC==names(mutation_list)[j]] <- MaxEntScan_diff_temp[which.max(abs(MaxEntScan_diff_temp))]}
    
  homozygous_edits_temp <- df_VEP_temp[xx]
    if (length(homozygous_edits_temp) >1){
      homozygous_edits_temp <- sort(homozygous_edits_temp)
      v <- homozygous_edits_temp[1]
      for (k in 2:length(homozygous_edits_temp)){
        v <- paste0(v,";",homozygous_edits_temp[k])
      }
       homozygous_edits_temp <- v
    }
    df_GT_puro_iBAR$homozygous_edits[df_GT_puro_iBAR$puroBC==names(mutation_list)[j]] <- homozygous_edits_temp
    
  }
}

df_GT_puro_iBAR$worst_consequence <- df_GT_puro_iBAR$consequence
df_GT_puro_iBAR$worst_consequence[grepl("stop codon",df_GT_puro_iBAR$worst_consequence)] <- "stop codon"
df_GT_puro_iBAR$worst_consequence[grepl("start lost",df_GT_puro_iBAR$worst_consequence)] <- "start lost"
df_GT_puro_iBAR$worst_consequence[grepl("splice variant",df_GT_puro_iBAR$worst_consequence)] <- "splice variant"
df_GT_puro_iBAR$worst_consequence[grepl("missense",df_GT_puro_iBAR$worst_consequence)] <- "missense"
df_GT_puro_iBAR$worst_consequence[grepl("promoter",df_GT_puro_iBAR$worst_consequence)] <- "promoter"
df_GT_puro_iBAR$worst_consequence[grepl("intron",df_GT_puro_iBAR$worst_consequence)] <- "intron"

df_GT_puro_iBAR$impact[grepl("HIGH",df_GT_puro_iBAR$impact)] <- "HIGH"
df_GT_puro_iBAR$impact[grepl("MODERATE",df_GT_puro_iBAR$impact)] <- "MODERATE"
df_GT_puro_iBAR$impact[grepl("MODIFIER",df_GT_puro_iBAR$impact)] <- "MODIFIER"

df_GT_puro_iBAR$PolyPhen[grepl("probably_damaging",df_GT_puro_iBAR$PolyPhen)] <- "probably_damaging"
df_GT_puro_iBAR$PolyPhen[grepl("possibly_damaging",df_GT_puro_iBAR$PolyPhen)] <- "possibly_damaging"
df_GT_puro_iBAR$PolyPhen[grepl("benign",df_GT_puro_iBAR$PolyPhen)] <- "benign"
df_GT_puro_iBAR$PolyPhen[grepl("-",df_GT_puro_iBAR$PolyPhen)] <- NA

df_GT_puro_iBAR$SIFT[grepl("deleterious",df_GT_puro_iBAR$SIFT)] <- "deleterious"
df_GT_puro_iBAR$SIFT[grepl("tolerated",df_GT_puro_iBAR$SIFT)] <- "tolerated"
df_GT_puro_iBAR$SIFT[grepl("-",df_GT_puro_iBAR$SIFT)] <- NA
```

Adding scores from a FACS-based screen from Coelho et al.(2023).

```{r}
original_gRNAs_info <- read.table("mutation_file_Coelho_2023.csv",sep=",",header=TRUE)#publicly available meta data from Coelho et al. (2023). Cell.
original_gRNAs_info <- original_gRNAs_info[grepl("BE3-NGG",original_gRNAs_info$editor),]
original_gRNAs_info_JAK1 <- original_gRNAs_info[grepl("JAK1",original_gRNAs_info$editor),]
intersected_gRNA <- intersect(original_gRNAs_info$sgRNA_ID,df_GT_puro_iBAR$gRNA)
intersected_gRNA_JAK1 <- intersect(intersected_gRNA,original_gRNAs_info_JAK1$sgRNA_ID)
df_GT_puro_iBAR$zscore_proliferation <- NA
for (j in 1:length(intersected_gRNA_JAK1)){
  df_GT_puro_iBAR$zscore_proliferation[df_GT_puro_iBAR$gRNA %in% intersected_gRNA_JAK1[j]] <- original_gRNAs_info_JAK1$zscore_proliferation[original_gRNAs_info_JAK1$sgRNA_ID == intersected_gRNA_JAK1[j]]
}
```

Adding mutational consequences based not on the actual mutations, but on those predicted to occur in Coelho et al.(2023).

```{r}
df_GT_puro_iBAR$predicted_consequence <- NA
df_GT_puro_iBAR$predicted_consequence[df_GT_puro_iBAR$gRNA %in% setdiff(df_GT_puro_iBAR$gRNA,intersected_gRNA)] <- "not predicted"

for (j in 1:length(intersected_gRNA)){
  df_GT_puro_iBAR$predicted_consequence[df_GT_puro_iBAR$gRNA == intersected_gRNA[j]] <- unique(original_gRNAs_info$Consequence[original_gRNAs_info$sgRNA_ID == intersected_gRNA[j]])
}
df_GT_puro_iBAR$predicted_consequence[!(is.na(df_GT_puro_iBAR$predicted_consequence)) & as.double(df_GT_puro_iBAR$start) >= 64964978 &  as.double(df_GT_puro_iBAR$start) <= 64967543] <- "promoter"


df_GT_puro_iBAR$citation <- NA
for (j in 1:length(intersected_gRNA)){
  df_GT_puro_iBAR$citation[df_GT_puro_iBAR$gRNA %in% intersected_gRNA[j]] <- unique(original_gRNAs_info$Citation[original_gRNAs_info$sgRNA_ID == intersected_gRNA[j]])
}
  
saveRDS(df_GT_puro_iBAR,file= "GT_barcode_metadata.rds")
```

Computing edit points


```{r}
compute_editing_position <- function(start,strand,GT){
  if (nchar(GT) >0 ){
    temp_GT <- strsplit(GT,";")[[1]]
    zygosity <- sapply(temp_GT,function(x) return(strsplit(x,"-")[[1]][2]))
    position <- as.double(sapply(temp_GT,function(x) return(strsplit(x,"-")[[1]][1])))
    if (strand == "+"){
      edit_position <- position + 1 - as.double(start)
    }else{
      edit_position <- as.double(start) + 23 - position
    }
    return(data.frame(edit_position=edit_position,zygosity=zygosity))
  }
}
df_GT_puro_iBAR_edited <- df_GT_puro_iBAR[df_GT_puro_iBAR$GT!="WT",]
editing_matrix <- matrix(ncol=length(-10:25),nrow=nrow(df_GT_puro_iBAR_edited))
colnames(editing_matrix) <- sapply(-10:25,toString)
rownames(editing_matrix) <- df_GT_puro_iBAR_edited$puroBC
edit_positions <- rep("",length(df_GT_puro_iBAR_edited$puroBC))
for (j in 1:nrow(editing_matrix)){
  tmp <- compute_editing_position(GT=df_GT_puro_iBAR_edited$GT[j],start=df_GT_puro_iBAR_edited$start[j],
                         strand=df_GT_puro_iBAR_edited$strand[j])
  tmp2 <- paste0(tmp$edit_position,"(",tmp$zygosity,")")
  v <- tmp2[1]
  if (length(tmp2) >1){
    for (k in 2:length(tmp2)){
      v <- paste0(v,";",tmp2[k])
    }
  }
  edit_positions[j] <- v
  xx <- intersect(colnames(editing_matrix),strsplit(toString(tmp$edit_position),", ")[[1]])
  if (length(xx) >0 ){
     editing_matrix[j,xx] <- tmp$zygosity[match(xx,strsplit(toString(tmp$edit_position),", ")[[1]])]
  }

}
df_GT_puro_iBAR_edited$editing_positions <- edit_positions
df_GT_puro_iBAR$editing_positions <- NA
df_GT_puro_iBAR[df_GT_puro_iBAR$GT!="WT",] <- df_GT_puro_iBAR_edited

editing_matrix_melt <- reshape2::melt(editing_matrix)
editing_matrix_melt <- editing_matrix_melt[complete.cases(editing_matrix_melt),]
colnames(editing_matrix_melt) <- c("barcode","position","zygosity")
ggplot(editing_matrix_melt,aes(x=barcode,y=position,fill=zygosity)) + geom_tile()+
 scale_fill_manual(values=c("0/0/1"= "darkblue","0/1/1"="purple","1/1/1"="darkred"),na.value = "white")+
  theme_classic() + geom_hline(yintercept=4) + geom_hline(yintercept=9)+ theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) 

df_GT_puro_iBAR
```


From bulk exome sequencing find mutations within the amplicon windows for comparison to frequent mutations found in the scDNA-seq data. 

```{r}
amplicons <- read.table("amplicons.txt")
lower_bound <- as.double(amplicons$V2)
upper_bound <- as.double(amplicons$V3)
library(vcfR)
exome_seq <- read.vcfR("HT-29.v1.caveman_c.annot.vcf")
exome_seq_chr1 <- exome_seq@fix[exome_seq@fix[,1]=="chr1" ,]
exome_gt_chr1 <- exome_seq@gt[exome_seq@fix[,1]=="chr1" ,]
exome_seq_chr1_positions <- as.double(exome_seq_chr1[,2])
xx <- list()
for (j in 1:length(lower_bound)){
  xx[[j]] <- exome_seq_chr1_positions  >= lower_bound[j]
  xx[[j]][exome_seq_chr1_positions > upper_bound[j]] <- FALSE
}
bulk_mutations <- list()
bulk_af <- list()
for (j in 1:length(lower_bound)){
  bulk_mutations[[j]] <- c()
  bulk_af[[j]] <- c()
  temp <- exome_seq_chr1[xx[[j]],,drop=FALSE]
  if (length(temp) > 0){
    temp_gt <- exome_gt_chr1[xx[[j]],,drop=FALSE]
    bulk_mutations[[j]] <- paste0(temp[1,2],"-",temp[1,4],"-",temp[1,5])
    bulk_af[[j]] <- strsplit(temp_gt[1,3],":")[[1]][10]
    if(length(temp) > 8){
      for (k in 1:nrow(temp)){
      bulk_mutations[[j]] <- c(bulk_mutations[[j]],paste0(temp[k,2],"-",temp[k,4],"-",temp[k,5]))
      bulk_af[[j]] <- c(bulk_af[[j]],strsplit(temp_gt[k,3],":")[[1]][10])
    }
    }
  }
}
df_bulk <- as.data.frame(cbind(unlist(bulk_mutations),unlist(bulk_af),(1:length(bulk_mutations))[sapply(bulk_mutations,function(x) !(is.null(x)))]))
colnames(df_bulk) <- c("mutations","af","amplicon")
df_bulk[,2]  <- as.double(df_bulk[,2])
df_bulk
saveRDS(df_GT_puro_iBAR,file= "GT_barcode_metadata.rds")
write.table(df_GT_puro_iBAR,"GT_barcode_metadata.csv",sep=",",col.names=TRUE,row.names=FALSE,quote=FALSE)
```

